---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
pkgload::load_all()
```


# RSQLCipher

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)

[![R-CMD-check](https://github.com/bpvgoncalves/RSQLCipher/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/bpvgoncalves/RSQLCipher/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/bpvgoncalves/RSQLCipher/branch/main/graph/badge.svg)](https://app.codecov.io/gh/bpvgoncalves/RSQLCipher?branch=main)
<!-- badges: end -->

Embeds the SQLCipher database engine in R and provides an interface compliant with the DBI package. 
[SQLCipher](https://www.zetetic.net/sqlcipher/open-source/) is an open source library that provides 
transparent and secure 256-bit AES encryption of SQLite database files. [SQLite](https://www.sqlite.org/index.html) 
is a public-domain, single-user, very light-weight database engine that implements a decent subset 
of the SQL 92 standard, including the core table creation, updating, insertion, and selection
operations, plus transaction management.  

This project started as a fork from the amazing [RSQLite](https://rsqlite.r-dbi.org) package for 
which its authors deserve full credit. It is intended to keep this package updated with new updates
to the RSQLite package to keep both as much compatible as possible.

This package allows for the use of regular SQLite database files or encrypted ones.


<!-- You can install the latest released version from CRAN with: -->

<!-- ```R -->
<!-- install.packages("RSQLCipher") -->
<!-- ``` -->

You can install the latest development version from GitHub with:

```R
# install.packages("devtools")
devtools::install_github("bpvgoncalves/RSQLCipher")
```

Discussions associated with DBI and related database packages take place on [R-SIG-DB](https://stat.ethz.ch/mailman/listinfo/r-sig-db).
The website [Databases using R](https://solutions.posit.co/connections/db/) describes the tools and 
best practices in this ecosystem.

## Usage

This package has identical interface to RSQLite.

```{r}
library(DBI)

# Create an two temporary RSQLCipher databases on disk.
# A regular database and an encrypted one.
tmp_plain <- tempfile()
con_plain <- dbConnect(RSQLCipher::SQLCipher(), 
                       tmp_plain)

tmp_enc <- tempfile()
con_enc <- dbConnect(RSQLCipher::SQLCipher(), 
                     tmp_enc, 
                     key = "0123456789ABCDEF0123456789ABCDEF0123456789ABCDEF0123456789ABCDEF") 

# Both databases behave the same on regular usage...
dbWriteTable(con_plain, "mtcars", mtcars)
dbWriteTable(con_enc, "mtcars", mtcars)

stopifnot(identical(dbListTables(con_plain), 
                    dbListTables(con_enc)))
stopifnot(identical(dbListFields(con_plain, "mtcars"), 
                    dbListFields(con_enc, "mtcars")))
stopifnot(identical(dbReadTable(con_plain, "mtcars"), 
                    dbReadTable(con_enc, "mtcars")))

dbDisconnect(con_plain)
dbDisconnect(con_enc)

# ... but the database files on disk are different
if ("hexView" %in% installed.packages()) {
  cat(">> Plain database\n")
  print(hexView::readRaw(tmp_plain, nbytes = 128))
  cat(">> Encrypted database\n")
  print(hexView::readRaw(tmp_enc, nbytes = 128))
} else {
  cat(">> Plain database\n")
  print(suppressWarnings(readLines(tmp_plain)[1:10]))
  cat(">> Encrypted database\n")
  print(suppressWarnings(readLines(tmp_enc)[1:10]))
}
file.remove(tmp_plain)
file.remove(tmp_enc)

```

